TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-13 paper/draft.tex
clarify gravity

We note that surface gravity is not inclided as a training label. The reason for this is that for main-sequence M dwarfs, the parameter is almost entirely redundant with metallicity. Unlike their higher-mass counterparts, M dwarf properties do not change measurably over the age of the Universe. Hence perfect knowledge of abundances and \teff\ for an M dwarf should uniquely determine its position on a color-magnitude diagram as well as its other fundamental properties. While we only have [Fe/H], for the uncertainties considered here, lack of information about other abundances will only be important for stars with extreme abundance patters (e.g., Carbon stars). Since we have no such targets in the training set, we could not handle such objects.

>>>>

We note that surface gravity is not inclided as a training label. The reason for this is that for main-sequence M dwarfs, the parameter is almost entirely redundant with metallicity. Unlike their higher-mass counterparts, M dwarf properties do not change measurably over the age of the Universe. Hence perfect knowledge of abundances and \teff\ for an M dwarf should uniquely determine its position its surface gravity, position on a color-magnitude diagram, and overall luminosity. While we only had [Fe/H] for the training sample, for the uncertainties considered here, lack of information about [$\alpha$/Fe] or specific abundances will only be important for extreme cases (e.g., Carbon stars). Highly active stars
